---
title: "EigenR"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

Originally, I entitled this package *Fast Matrix Algebra with 'Eigen'*, because 
I expected it to be faster than R base. But this is not the case. So I entitled 
it *Complex Matrix Algebra with 'Eigen'*, because it supports some operations 
on complex matrices which are not supported by R base: determinant, Cholesky 
decomposition, and linear least-squares problems.

```{r packages}
library(EigenR)
library(microbenchmark)
```

## Determinant

```{r det_benchmark}
set.seed(666L)
M <- matrix(rnorm(300L*300L, mean = 1), 300L, 300L)
M[sample.int(300L*300L, 300L*270L)] <- 0 # 90% of zeros
Ms <- asSparseMatrix(M)
microbenchmark(
  base          = det(M),
  EigenR        = Eigen_det(M),
  EigenR_sparse = Eigen_det(Ms), # :-(
  times = 200L
)
```

Determinants of complex matrices are supported:

```{r det_cplx_benchmark}
set.seed(666L)
Mr <- matrix(rnorm(100L*100L, mean = 1), 100L, 100L)
Mi <- matrix(rnorm(100L*100L, mean = 1), 100L, 100L)
M <- Mr + 1i * Mi
library(complexplus)
microbenchmark(
  EigenR      = Eigen_det(M), # :-)
  complexplus = Det(M), 
  times = 30L
)
```

## Inverse matrix

```{r inverse_benchmark}
set.seed(666L)
M <- matrix(rnorm(100L*100L), 100L, 100L)
microbenchmark(
  base   = solve(M),
  EigenR = Eigen_inverse(M), # :-(
  times = 500L
)
```

## Cholesky decomposition

```{r chol_benchmark}
set.seed(666L)
M <- matrix(rpois(10000L, 25), 100L, 100L)
A <- crossprod(M)
microbenchmark(
  base   = chol(A),
  EigenR = Eigen_chol(A), # :-(
  times = 1000L
)
```

Cholesky decomposition of complex matrices is supported.

## Pivoted Cholesky decomposition

```{r UtDU_benchmark, warning=FALSE}
set.seed(666L)
M <- matrix(rgamma(202L*199L, 10), 202L, 199L)
M <- cbind(M, M[, 1L] + 3*M[, 2L])
A <- crossprod(M)
microbenchmark(
  base   = chol(A, pivot = TRUE),
  EigenR = Eigen_UtDU(A), # :-(
  times = 1000L
)
```

Pivoted Cholesky decomposition of complex matrices is supported.

## Kernel

```{r kernel_benchmark}
set.seed(666L)
M <- matrix(rpois(10000L, 25), 100L, 100L)
A <- cbind(M, M)
At <- t(A)
library(MASS)
microbenchmark(
  MASS       = Null(At),
  EigenR_LU  = Eigen_kernel(A, method = "LU"),  # :-)
  EigenR_COD = Eigen_kernel(A, method = "COD"), 
  times = 100L
)
```

## Linear least-squares problems

```{r lsSolve_benchmark}
set.seed(666L)
n <- 700L; p <- 200L
A <- matrix(rnorm(n * p), n, p)
b <- rnorm(n)
microbenchmark(
  stats   = lm.fit(A, b),
  Eigen_R = Eigen_lsSolve(A, b), # :-(
  times = 20L
)
```

Complex matrices `A` and `b` are supported.
